The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2011-11-28 smurf_paper.tex
Complete steps section in main body of paper
The appendix though is not yet complete.

Steps occur with a wide range of heights and shapes. The ratio of step
height to noise can vary from less than 10 to several hundred. Some steps
occur over a single sample such as the step close to sample 5000 in the
right-hand plot of Fig.~\ref{fig:steps1}, but others happen more
gradually such as the step close to sample 5300. In addition, a step can
be preceeded or followed by a short period of instability, as is visible
at the bottom of the step in the left hand plot of
Fig.~\ref{fig:steps1}. Further problems are caused by steps that occur
close together in time, such as the large downward step followed by a
smaller upward step close to sample 5000 in the right hand plot of
Fig.~\ref{fig:steps1}.

Detecting and correcting such a wide variety of steps reliably has proved
to be a challenge. The current step-correction algorithm is described in
detail in appendix~\ref{appendix:steps}. In outline, the following stages
are involved in detecting and correcting steps in a single bolometer time
stream:

@@

\item Median smooth the whole time stream.
\item Find the gradient of the median smoothed time series at each sample.
\item Smooth the gradient values to determine the local mean gradient and
subtract this local mean from the total gradient to get residual gradients.
\item Find residual gradient values that exceed 25 times the local RMS of
the residual gradients.
\item Group these high residual gradients into contiguous blocks of
samples.
\item Merge blocks that are separated by less than 100 samples

@@

>>>>

Steps occur with a wide range of heights and shapes. The ratio of step
height to noise can vary from less than 10 to several hundred. Some steps
occur over a single sample such as the step close to sample 5000 in the
right-hand plot of Fig.~\ref{fig:steps1}, but others happen more
gradually such as the step close to sample 5300. In addition, a step can
be preceded or followed by a short period of instability, as is visible
at the bottom of the step in the left hand plot of
Fig.~\ref{fig:steps1}. Further problems are caused by steps that occur
close together in time, such as the large downward step followed by a
smaller upward step close to sample 5000 in the right hand plot of
Fig.~\ref{fig:steps1}.

Detecting and correcting such a wide variety of steps reliably has proved
to be a challenge. The current step-correction algorithm is described in
detail in appendix~\ref{appendix:steps}. In outline, the following stages
are involved in detecting steps in a single bolometer time stream:

\begin{enumerate}
\item Median smooth the whole time stream.
\item Find the gradient of the median smoothed time series at each sample.
\item Smooth the gradient values to determine the local mean gradient and
subtract this local mean from the total gradient to get residual gradients.
\item Find residual gradient values that exceed 25 times the local RMS of
the residual gradients.
\item Group these high residual gradients into contiguous blocks of
samples.
\item Merge blocks that are separated by less than 100 samples
\end{enumerate}

The above process produces a list of candidate steps in each bolometer
time series. Each candidate step is then verified, measured and corrected
using the following procedure:

@@

\item The above process can mis-interpret the edges of a bright source
as a step. So we ignore blocks that occur close to bright sources.

\item If the block passes the above test, a least squares linear fit is
performed to the median-filtered bolometer data just before the block, and
this fit is extrapolated to predict the data value expected at the centre
of the block on the basis of the preceding data.

\item A least squares linear fit is performed to the median-filtered
bolometer data just after the block, and this fit is extrapolated to
predict the data value expected at the centre of the block on the basis
of the following data.

\item The difference between these two expected data values is taken as
the step height.

\item The preceding three steps are repeated several times, each time
including a different selection of samples in the two least squares fits.
The mean and standard deviation of the corresponding set of step heights
is found.

\item If the mean step height is small compared to the standard deviation
of the step heights, or compared to the noise in the bolometer data, then
the step is ignored.

\item If the above checks are passed, all subsequent bolometer samples are
corrected by the mean step height.

\item Bolometer samples within the duration of the step, and a few
samples on either side, are flagged as unusable.

@@

Once all steps have been corrected within a bolometer time stream, a
constant value is added to all samples in the time stream to restore its
original mean value.
